---
output: 
  github_document:
    html_preview: false
always_allow_html: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  fig.align = "center",
  comment = "#>",
  message = FALSE,
  warning = FALSE
)
```

# COREnets

## Installation

  1. Install `devtools` if you haven't already.

```{r, eval=FALSE}
install.packages("devtools")
```

  2. Install the package using `devtools`.
  
```{r, eval=FALSE}
devtools::install_github("NPSCORELAB/COREnets")
```

## Using the Package

`COREnets` contains a series of network datasets that can be accessed using explicit namespace access:

```{r}
library(COREnets)
anabaptists <- COREnets::anabaptists
```

Each data object is a nested list with three levels:

```{r}
names(anabaptists)
```

  1. The `metadata` contains information on the dataset:
```{r, echo=F}
library(tidyverse)
metadata <- tibble::tibble(
  Field = c(
    "title",
    "name",
    "category",
    "tags",
    "description"
  ),
  Type = c(
    "`character",
    "`character",
    "`character",
    "`character",
    "`character"
  ),
  Definition = c(
    "A formal title for the dataset for external uses.",
    "An informal dataset label for internal use.",
    "An internal classification for the type dataset, based on one of the following: <br>
    - Religious <br>
    - Terrorirsm <br>
    - Criminal <br>
    - Other",
    "An atomic vector of key words assinged to the piece of data.",
    "A brief definition of the dataset in regards to the type of data, collection, etc."
  )
)
metadata %>%
  knitr::kable(format = "html", escape = FALSE) %>%
  kableExtra::kable_styling(bootstrap_options = c("bordered"))
```
  
  2. The `bibtex_data` the data fields required to generate a [bibtex citation](https://verbosus.com/bibtex-style-examples.html). Note that the some datasets will have mutiple citation entries.
  3. The `network` field contains a list of meta data and both the node and edges tables required to generate a network graph:
  
  - `net_metadata`: A list of descriptive information on the type of network:
```{r, echo=FALSE}
net_metadata <- tibble::tibble(
  Field = c(
    "node_type",
    "edge_type",
    "modes",
    "directed",
    "weighted",
    "multiplex",
    "dynamic",
    "spatial"
  ),
  Type = c(
    "character",
    "character",
    "double",
    "logical",
    "logical",
    "logical",
    "logical",
    "logical"
  ),
  Definition = c(
    "Description of the node class.",
    "Description of the edge.",
    "A number denoting one or two mode networks.",
    "A logical denoting whether the network edges are directed or not.",
    "A logical denoting whether or not the edges are weighted.",
    "A logical denoting whether or not the network is made up of multiple layers.",
    "A logical denoting whether or not the network changes with time.",
    "	A logical denoting whether or not the network vertices or edges are associated with a geographic feature."
  )
)

net_metadata %>%
  knitr::kable(format = "html", escape = FALSE) %>%
  kableExtra::kable_styling(bootstrap_options = c("bordered"))
```
  
  - `node_table`: A `data.frame` containg node attributes. A unique identifier for each node in the `edge_table` should be present in the `id` variable. 
  - `edge_table`: A `data.frame` that contains a minimum of two columns, one column of nodes acting as a vector source or starting point (`from`) and another column of nodes that are the target of the connection (`to`).